### 2014-10-21 thesis/method/classification.tex
More writing about classification algorithms...

\chapter{LOS Classification and Supervised Learning}
  \label{chap:classification}

@@

1R works by considering each feature individually.
For each of the values of the
feature, 1R finds the class with the highest frequency and creates a rule
from this observation. If the feature is numeric (that is, its values are
continuous rather than discrete), then 1R will partition the spectrum of
values into discrete buckets. This is known as \textit{discretisation}.
The error rate, expressed as the ratio of correct predictions to the total
number of examples in the data set,
is calculated for each rule that is generated
in this way. The rule with the lowest error rate is then chosen as the 1-rule
for this classifier,
and if there is a tie, it is broken randomly. As with 0R, the prediction
decision of 1R is transparent and it provides another baseline upon which
more sophisticated classifiers can be evaluated.

@@

For a class variable with two possible values that occur with
probability $p$ and $1-p$ respectively, and a feature vector
$(F_1,F_2,\ldots,F_n)$, the logistic regression model is:
\begin{equation*}
\mathrm{logit}(p) &=& \mathrm{log} \left(\dfrac{p}{1-p}\right) \\
  &=& \beta_0 + \beta_1F_1 + \beta_2F_2 + \ldots + \beta_nF_n
\end{equation*}
This can be rearranged to:
\begin{equation*}
p &=& \mathbb{P}(Class=\mathrm{LOS }>2\mathrm{ days}|F_1,F_2,\ldots,F_n) \\
  &=& \dfrac{e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}{1 + e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}
\end{equation*}

@@

@@

\subsection{Extensions}
There have been many extensions proposed to the $k$-NN algorithm described
above.

\subsubsection{K*: an Entropy-Based Distance Function}

\subsubsection{Ranked Distance}

@@

>>>>

\chapter{LOS Classification Using Supervised Learning}
  \label{chap:classification}

@@

1R works by considering each feature individually.
For each of the values of the
feature, 1R finds the class with the highest frequency and creates a rule
from this observation. If the feature is numeric (that is, its values are
continuous rather than discrete), then 1R will partition the spectrum of
values into discrete buckets. This is known as \textit{discretisation}.
The error rate, expressed as the ratio of correct predictions to the total
number of examples in the data set,
is calculated for each rule that is generated
in this way. The rule with the lowest error rate is then chosen as the 1-rule
for this classifier,
and if there is a tie, it is broken randomly. As with 0R, the prediction
decision of 1R is transparent and it provides another baseline upon which
more sophisticated classifiers can be evaluated. Studies have also found that
1R performs quite well in terms of accuracy relative to other classifiers
\cite{Holte1993}.

@@

For a class variable with two possible values that occur with
probability $p$ and $1-p$ respectively, and a feature vector
$(F_1,F_2,\ldots,F_n)$, the logistic regression model is:
\begin{align*}
\mathrm{logit}(p) &= \mathrm{log} \left(\dfrac{p}{1-p}\right) \\
  &= \beta_0 + \beta_1F_1 + \beta_2F_2 + \ldots + \beta_nF_n
\end{align*}
For LOS classification, this can be rearranged to:
\begin{align*}
p &= \mathbb{P}(Class=\mathrm{LOS }\leq2\mathrm{ days}|F_1,F_2,\ldots,F_n) \\
  &= \dfrac{e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}{1 + e^{\beta_0 + \sum^n_{i=1}\beta_i F_i}}
\end{align*}

@@

To classify a feature vector $x = (x_1,x_2,\ldots,x_n)$, we substitute the
$x_i$'s into the above equation in place of the $F_i$'s, which gives us a
probability. If this probability is above a particular threshold, then we
classify the example as belonging to the class ``LOS $\leq$ 2 days'', otherwise
it is classified as ``LOS $>$ 2 days''. Usually this threshold is 0.5.

@@

\subsection{Extensions}
There have been many extensions proposed to the $k$-NN algorithm described
above. These usually seek to find a more suitable distance metric for the
problem domain at hand, or to reduce the storage requirements of the original
$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
K* and our contribution, Ranked Distance.

\subsubsection{K*: Instance-based Classifier with Entropic Distance Function}

\subsubsection{Ranked Distance}
\paragraph{Intuition}
As we mentioned earlier, all features are considered equally in a $k$-NN
classifier, which implies that all features are equally important in predicting
the class. In practice, however, this is not always the case: to predict the
LOS of a trauma patient, the severity of their injury should affect the final
LOS more than whether or not they can speak English (which could be just two
of many features that are recorded about a patient). Our contribution is thus
a modified distance function to take the relative importance of attributes into
account.

\paragraph{Mathematical formulation}
Given two instances $x = (x_1,x_2,\ldots,x_n)$ and $y = (y_1,y_2,\ldots,y_n)$,
the distance between them is given by:
\begin{equation*}
\mathrm{Distance} = \sum_{i=1}^n w_i |x_i-y_i|
\end{equation*}
where the $w_i$s weight the contribution of the $i$-th feature to the overall
distance that is computed by the $k$-NN algorithm. We assume that all feature
values have been normalised, as described above.

\paragraph{Assignment of weights}
The weights $w_i$ can be tuned to match the particular problem that is being
investigated. For our LOS classification problem, we will consider two ways
of weighting features:
\begin{enumerate}
\item The weight of each feature is the magnitude of the correlation
coefficient
between it and the class, meaning that features that are more highly correlated
with the class will contribute more to the distance. The correlation
coefficient $r_i$ for feature $i$ with the class $c$ can be computed by:
\begin{equation*}
r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
\end{equation*}
where $m$ is the number of training examples, $x_i$ is the value of feature
$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
values and the class values respectively. 
\item We compute the correlation coefficients as described in the point above.
This gives us a ranking of the importance of each feature to the prediction of
the class, with 1 being the highest rank (indicating the most importance) and
$n$ being the lowest rank (indicating the least importance).
Instead of using the values of the correlation directly to weight
the distance contribution of each feature, we specify a function:
\begin{equation*}
f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
\end{equation*}
that describes
how the weights of the features vary with the rank. This function should
inituitively be non-increasing and should produce lower values when the rank
number is greater, meaning that less important features have a lower weight
in the distance calculation.
Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
results in all features contributing equally to the distance calculation.

\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}}$ and
$f(\mathrm{Rank}) = n-\mathrm{Rank}$.
\end{enumerate}

@@

\section{Conclusion}
